---
title: "HW 04"
output: html_document
date: '2022-11-20'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# HPC

## Problem 1

```{r}
# Total row sums
fun1 <- function(mat) {
  n <- nrow(mat)
  ans <- double(n) 
  for (i in 1:n) {
    ans[i] <- sum(mat[i, ])
  }
  ans
}

fun1alt <- function(mat) {
  rowSums(mat)
}

# Cumulative sum by row
fun2 <- function(mat) {
  n <- nrow(mat)
  k <- ncol(mat)
  ans <- mat
  for (i in 1:n) {
    for (j in 2:k) {
      ans[i,j] <- mat[i, j] + ans[i, j - 1]
    }
  }
  ans
}

fun2alt <- function(mat) {
  t(apply(mat,1,cumsum))
}


# Use the data with this code
set.seed(2315)
dat <- matrix(rnorm(200 * 100), nrow = 200)

# Test for the first
microbenchmark::microbenchmark(
  fun1(dat),
  fun1alt(dat),  check = "equivalent"
)

# Test for the second
microbenchmark::microbenchmark(
  fun2(dat),
  fun2alt(dat), check = "equivalent"
)
```

# Problem 2

```{r}
sim_pi <- function(n = 1000, i = NULL) {
  p <- matrix(runif(n*2), ncol = 2)
  mean(rowSums(p^2) < 1) * 4
}

# Here is an example of the run
set.seed(156)
sim_pi(1000) # 3.132
```

```{r}
# This runs the simulation a 4,000 times, each with 10,000 points
set.seed(1231)
system.time({
  ans <- unlist(lapply(1:4000, sim_pi, n = 10000))
  print(mean(ans))
})
```

```{r}
library(parallel)

cl <- makeCluster(getOption("cl.cores", 2))
set.seed(1231)
RNGkind("L'Ecuyer-CMRG")
set.seed(1231)
(s <- .Random.seed)
clusterSetRNGStream(cl,s)
system.time({
  
  ans <- unlist(parLapply(cl, 1:4000, sim_pi, n = 10000))
  print(mean(ans))
})
```

# SQL
```{r}
# install.packages(c("RSQLite", "DBI"))

library(RSQLite)
library(DBI)

# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
film <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film.csv")
film_category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film_category.csv")
category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/category.csv")

# Copy data.frames to database
dbWriteTable(con, "film", film)
dbWriteTable(con, "film_category", film_category)
dbWriteTable(con, "category", category)
```

## Question 1
How many many movies is there available in each rating catagory.

```{sql, connection=con}
SELECT rating, COUNT(*) AS "movie_count"
FROM film
GROUP BY rating
```

## Question 2
What is the average replacement cost and rental rate for each rating category.
```{sql, connection=con}
SELECT rating, AVG(replacement_cost) AS "average of replacement cost", AVG(rental_rate) AS "average of rental rate"
FROM film
GROUP BY rating
```

##Question 3
Use table film_category together with film to find the how many films there are with each category ID
```{sql, connection=con}
SELECT category_id, COUNT(*) AS movie_count
FROM film
INNER JOIN film_category
ON film.film_id = film_category.film_id
GROUP BY category_id
```

## Question 4
Incorporate table category into the answer to the previous question to find the name of the most popular category.
```{sql, connection=con}
SELECT category.name, COUNT(*) AS movie_count
FROM film
INNER JOIN film_category
ON film.film_id = film_category.film_id
INNER JOIN category
ON film_category.category_id = category.category_id
GROUP BY category.name
ORDER BY movie_count DESC
```
According to the movie count, the most popular category is sports.

### Clean-up
```{r}
dbDisconnect(con)
```