---
title: "lab4"
author: "Eleanor Bi"
date: "10/5/2022"
output: html_document
---
```{r, echo=TRUE}
EVAL <- FALSE
ECHO <- EVAL
library(tidyverse)
library(lubridate)
library(data.table)
```

## 1. Read in the data
```{r, echo=TRUE}
if (!file.exists("met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )
met <- data.table::fread("met_all.gz")
```

## 2. Prepare the data
```{r}
met <- met[temp >= -17][elev == 9999.0, elev := NA]
summary(met$temp)
summary(met$elev)
summary(met$wind.sp)
#Generate a date variable using the functions as.Date() (hint: You will need the following to create a date paste(year, month, day, sep = "-")).
library(lubridate)
met= met[ , ymd := as.Date(paste(year, month, day, sep = "-"))]
met[, table(week(ymd))]
met= met[week(ymd)==31]
#Compute the mean by station of the variables temp, rh, wind.sp, vis.dist, dew.point, lat, lon, and elev.
met[, .(
  temp= max(temp, na.rm=T),
  rh= max(rh, na.rm=T),
  wind.sp= max(wind.sp, na.rm=T),
  vis.dist= max(vis.dist, na.rm=T), 
  dew.point= max(dew.point, na.rm=T), 
  lat= max(lat, na.rm=T), 
  lon= max(lon, na.rm=T), 
  elev= max(elev, na.rm=T)
)]
met_avg= met[, .(
  temp= mean(temp, na.rm=T),
  rh= mean(rh, na.rm=T),
  wind.sp= mean(wind.sp, na.rm=T),
  vis.dist= mean(vis.dist, na.rm=T), 
  dew.point= mean(dew.point, na.rm=T), 
  lat= mean(lat, na.rm=T), 
  lon= mean(lon, na.rm=T), 
  elev= mean(elev, na.rm=T)
), by= 'USAFID']
#create a region variable
library(dtplyr)
library(data.table)
met_avg= met_avg[ , region := fifelse(lon>= -98 & lat> 39.71, "NE",
                 fifelse(lon< -98 & lat> 39.71, "NW",
                 fifelse(lon< -98 & lat<= 39.71, "SW","SE")))
     ]
table(met_avg$region)
#create a categorical variable for elevation
met_avg[, elev_cat := fifelse(elev >252, "high", "low")]
```

## 3. Use geom_violin to examine the wind speed and dew point temperature by region
```{r}
library(ggplot2)
met_avg[!is.na(region)] %>% 
  ggplot() + 
  geom_violin(mapping = aes(x = 1, y = dew.point, color=region, fill= region)) + 
  facet_wrap(~ region, nrow = 1)
```
Thus, the highest dew point temp is in se.
```{r}
met_avg[!is.na(region) & !is.na(wind.sp)] %>% 
  ggplot() + 
  geom_violin(mapping = aes(x = 1, y = wind.sp, color=region, fill= region)) + 
  facet_wrap(~ region, nrow = 2)
```
Thus, ne has the highest wind speed

##4. Use geom_jitter with stat_smooth to examine the association between dew point temperature and wind speed by region
```{r}
met_avg[!is.na(region) & !is.na(wind.sp)] %>% 
  ggplot(mapping = aes(x = wind.sp, y = dew.point)) + 
  geom_point(mapping= aes(color= region)) + 
  geom_smooth(method= lm, mapping= aes(linetype= region))+
  facet_wrap(~ region, nrow = 2)
```

## 5. Use geom_bar to create barplots of the weather stations by elevation category coloured by region

## 6. Use stat_summary to examine mean dew point and wind speed by region with standard deviation error bars
```{r}
met_avg[!is.na(dew.point)] %>%
  ggplot(mapping = aes(x = region, y = dew.point)) + 
  stat_summary(fun.data= mean_sdl, geom= "errorbar")+
  stat_summary(fun.data= mean_sdl)
```
## 7. 
```{r}
library(leaflet)
rh.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=met_avg$rh)
rh.pal
top10rh= met_avg[rank(-rh) <=10][1:10]

```

## 8.Use a ggplot extension
```{r}
rhmap <- leaflet(met_avg) %>% 
  # The looks of the Map
  addProviderTiles('CartoDB.Positron') %>% 
  # Some circles
  addCircles(
    lat = ~lat, lng=~lon,
                                                  # HERE IS OUR PAL!
    label = ~paste0(rh), color = ~ rh.pal(rh),
    opacity = 1, fillOpacity = 1, radius = 500
    ) %>%
  # And a pretty legend
  addLegend('bottomleft', pal=rh.pal, values=met_avg$rh,
          title='Releative Hunid.', opacity=1)
rhmap
```

