---
title: "lab3"
author: "Eleanor Bi"
date: "10/5/2022"
output: html_document
---
## 1. Read in the data
```{r}
library('R.utils')
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", method="libcurl", timeout = 60)
met <- data.table::fread("met_all.gz")

```

## 2. Check the dimensions, headers, footers. How many columns, rows are there?
```{r}
dim(met)
head(met)
tail(met)
```
There are 2,377,343 rows and 30 columns in the met dataset.

## 3. Take a look at the variables
```{r}
str(met)
```

## 4. Take a closer look at the key variables
```{r}
table(met$year)
table(met$day)
table(met$hour)
summary(met$temp)
summary(met$elev)
summary(met$wind.sp)
#After checking the data we should make the appropriate modifications. Replace elevations with 9999 as NA.
met[met$elev==9999.0] <- NA
summary(met$elev)
```
At what elevation is the highest weather station?

The weather station with highest elevation is 4113 meters. This is after replacing 9999.0 values with the appropriate code for “missing”, which is “NA”.

```{r}
met <- met[temp>-40]
met2 <- met[order(temp)]
head(met2)
```

## 5. Check the data against an external data source.

```{r}
met <- met[temp>-15]
met2 <- met[order(temp)]
head(met2)
```

## 6.Calculate summary statistics
initial question!!:
We want to pick out the weather station with maximum elevation and examine its windspeed and temperature.
```{r}
elev <- met[elev==max(elev)]
summary(elev)
```
```{r}
cor(elev$temp, elev$wind.sp, use="complete")
cor(elev$temp, elev$hour, use="complete")
cor(elev$wind.sp, elev$day, use="complete")
cor(elev$wind.sp, elev$hour, use="complete")
cor(elev$temp, elev$day, use="complete")
```


## 7. Exploratory graphs
```{r}
hist(met$elev, breaks=100)
```
```{r}
hist(met$temp)
```
```{r}
hist(met$wind.sp)
```
```{r}
library('dplyr')
library(leaflet)
leaflet(elev) %>%
  addProviderTiles('OpenStreetMap') %>% 
  addCircles(lat=~lat,lng=~lon, opacity=1, fillOpacity=1, radius=100)
```

```{r}
library(lubridate)
elev$date <- with(elev, ymd_h(paste(year, month, day, hour, sep= ' ')))
summary(elev$date)
```

```{r}
elev <- elev[order(date)]
head(elev)
```
With the date-time variable we can plot the time series of temperature and wind speed.
```{r}
plot(elev$date, elev$temp, type='l')
```
```{r}
plot(elev$date, elev$wind.sp, type='l')
```


