---
title: "lab6"
author: "Eleanor Bi"
date: "12/13/2022"
output: html_document
---
```{r}
library(webshot)
webshot::install_phantomjs()
```

```{r read-data, cache= TRUE}
if (!file.exists("mtsamples.csv")) {
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/00_mtsamples/mtsamples.csv",
    "mtsamples.csv", method = "libcurl", timeout  = 60)
}
mts <- read.csv("mtsamples.csv")
str(mts)
mts= as_tibble(mts)
mts
```
## Question 1: What specialties do we have?
We can use count() from dplyr to figure out how many different catagories do we have? Are these catagories related? overlapping? evenly distributed?
```{r mdeical_specialties}
specialties=
  mts %>%
  count(medical_specialty)
specialties %>%
  arrange(desc(n)) %>%
knitr::kable()
```
```{r barplot-of-specialty-counts}
specialties %>%
  top_n(10) %>%
  ggplot( aes(x= n, y= fct_reorder(medical_specialty, n))) +
  geom_col()
```
they are not all uniform.

## Question 2
Explain what we see from this result. Does it makes sense? What insights (if any) do we get?
```{r token-transcription, cache=TRUE}
mts %>%
  unnest_tokens(word, transcription) %>%
  #anti_join(stop_words, by = c("word")) %>%
  count(word, sort = TRUE) %>%
  top_n(20, n) %>%
  ggplot(aes(n, fct_reorder(word, n))) +
  geom_col()
```
We can found that there are a lot of stop words here, non-specific to medical text.

## Question 3: Redo analysis in Q2, amd remove stopwords
```{r, cache=TRUE}
mts %>%
  unnest_tokens(word, transcription) %>%
  count(word, sort = TRUE) %>%
  anti_join(stop_words, by = c("word")) %>%
  # use regular expression to filter out numbers (not match 0-9)
  filter( !grepl(pattern= "^[0-9]+$", x= word)) %>%
  top_n(20, n) %>%
  ggplot(aes(n, fct_reorder(word, n))) +
  geom_col()
```

## Question 4: repeat question 2, but this time tokenize into bi-grams. how does the result change if you look at tri-grams?
```{r bigrams-transcription, cache=TRUE}
mts %>%
  unnest_ngrams(bigram, transcription, n=2) %>%
  count(bigram, sort = TRUE) %>%
  top_n(20, n) %>%
  ggplot(aes(n, fct_reorder(bigram, n))) +
  geom_col()
```
```{r trigrams-transcription, cache=TRUE}
mts %>%
  unnest_ngrams(trigram, transcription, n=3) %>%
  count(trigram, sort = TRUE) %>%
  top_n(20, n) %>%
  ggplot(aes(n, fct_reorder(trigram, n))) +
  geom_col()
```
The top 20 trigrams seemed to find a few more medical word group than bigrams.

## Question 5
Using the results you got from questions 4. Pick a word and count the words that appears after and before it.
```{r bigrams-transcription-nextword, cache=TRUE}
ptbigram =
  mts %>%
  unnest_ngrams(bigram, transcription, n=2) %>%
  separate(bigram, into = c("word1", "word2"), sep = " ") %>%
  select(word1, word2) %>%
  filter(word1 == "patient" | word2 == "patient")
```
Words appearing before patient:
```{r before-patient} 
ptbigram %>%
  filter(word2 == "patient") %>%
  count(word1, sort=T) %>%
  anti_join(stop_words, by= c("word1"= "word")) %>%
  top_n(10) %>%
knitr::kable()
```
Words appearing after patient:
```{r after-patient} 
ptbigram %>%
  filter(word1 == "patient") %>%
  count(word2, sort=T) %>%
  anti_join(stop_words, by= c("word2"= "word")) %>%
  top_n(10) %>%
knitr::kable()
```




